---
title: "OneSite_StormSelection"
author: "Marc Peipoch"
date: '2023-01-24'
output: html_document
---

This code allows visual inspection of the c-Q relationships for each selected storm. I wrote down the storm_ID of those showing clear
patterns of hysteresis loops and validated they were indeed storm vents (signifcant increase in Q) by axomining the cha_AllCQData file when necessary
```{r setup, include=FALSE}
library(ggplot2) ; library(dplyr) ; detach("package:tidyr", unload=TRUE); library(broom)

site="BLUE"

setwd(paste("R:/EcosystemEcologyLab/MSAPLANKTONdataDirectory/DataInDevelopment/NEONData/",site,"/04_product",sep=""))
Storms_2017 = read.csv("Chlorophyll_2017/chl_AllCQData.csv", header = T)
Storms_2018 = read.csv("Chlorophyll_2018/chl_AllCQData.csv", header = T)
Storms_2019 = read.csv("Chlorophyll_2019/chl_AllCQData.csv", header = T)
Storms_2020 = read.csv("Chlorophyll_2020/chl_AllCQData.csv", header = T)
Storms_2021 = read.csv("Chlorophyll_2021/chl_AllCQData.csv", header = T)
Storms_2022 = read.csv("Chlorophyll_2022/chl_AllCQData.csv", header = T)

```
```{r}
plot_data = Storms_2019


storm_numb = unique(plot_data$storm_id)
plot_results = list()
  
for (i in 1:length(storm_numb)) {
  plot_subdata = subset(plot_data, storm_id==storm_numb[i])
  print(ggplot(plot_subdata, aes(q_cms, conc, colour=time_step, size=1)) + 
  geom_point() + scale_colour_gradient(low = "yellow", high = "red")
 + ggtitle(print(storm_numb[i])))
  
}


```



This code corrects the mismatch in storm_ID dates that I discovered while compiling the chl_HysteresisData results for the selected events. The dates and storm_ID in the cha_AllCQData file are the one that matched the raw data and the ones I used above for visual inspection. 

Therefore, the following code will use the datetime per storm_ID from the the cha_AllCQData file to:

1) Calculate Chl export
2) calculate the total increase in [chl] concentration 
3) calculate the C-Q slope
4) link 1 and 2 to each storm ID event
```{r}
####################parameters to adjust
site = "TOMB"
year = "2021"
parameter="Chlorophyll_"
selected_storms = paste0('storm_',c(3))
###########################################

cQ_file = "chl_AllCQData.csv"
hyst_file= "chl_HysteresisData.csv"
####################parameters to adjust



setwd(paste("R:/EcosystemEcologyLab/MSAPLANKTONdataDirectory/DataInDevelopment/NEONData/",site,"/04_product/",parameter,year,sep=""))

AllCQData = read.csv(cQ_file, header = T)
 

#calculate the export for chlorophyll per event
exportChl = AllCQData %>%
  group_by(storm_id) %>%
  mutate(risignChlexport = conc*(q_cms*1000)*(60*30)) %>% #in ug Chl every 30min
  summarize(totalChlexport = sum(risignChlexport)/1000000) #in g Chl per event

AllCQData_v2 = left_join(AllCQData,exportChl,by="storm_id")

#calculate the differential increase/decrease in Chl concentration per storm event 
increaseChl = AllCQData %>%
  group_by(storm_id) %>%
  slice(1:which.max(conc)) %>%
  summarize(deltaChl = max(conc) - min(conc)) 

AllCQData_v3 = left_join(AllCQData_v2,increaseChl,by="storm_id")


#calculate the the C-Q slope
CQslope = AllCQData %>%
  group_by(storm_id) %>%
  do(slope = as.numeric(coef(lm(log(conc+0.01) ~ log(q_cms),data=.))[2]))

CQslope = as.data.frame(CQslope)
AllCQData_v4 = left_join(AllCQData_v3,CQslope,by="storm_id")


#first extract the unique set of dates we need 

subset_CQData = AllCQData_v4 %>%
  filter(., storm_id %in% selected_storms) %>%
  mutate(start=datetime) %>%
  dplyr::select(start,totalChlexport,deltaChl,slope) 
    
    
#input hysteresis indices results and join by the selected dates

AllHysteresisData = read.csv(hyst_file, header = T)


temp = left_join(subset_CQData,AllHysteresisData,by="start")
temp = temp[!is.na(temp$storm_id),]
temp = apply(temp,2,as.character)

file.path = "R:/EcosystemEcologyLab/MSAPLANKTONdataDirectory/DataInDevelopment/Aggregations/AggregatedNEONdata/01_input/Chlorophyll_export&calcs/"
write.csv(temp,file = file.path(file.path,paste(site,year,parameter,"Results.csv",sep="_")))

print(selected_storms)

```
